### 2021-06-04 text/article_vedrovice.Rmd
cite packages

The given statistics for the independent random process are estimated using Monte Carlo
simulation with 999 iterations.
Although a smaller number of permutations should be sufficient to achieve
credible results
[@baddeley2016, 384], we follow the approach taken by @sayer2013 [78].
The ten most extreme realizations of the simulation are discarded as outliers, so
the significance level is 0.01 [cf. @baddeley2016 410].
The remaining realizations of given point processes are used to construct envelopes
around function curves derived from observed data.

@@

<!-- software packages -->
Data is analyzed in an R environment for statistical computing and graphics
[@rcoreteam2021] using functions of several different packages; for the complete list,
see the supplementary information.

@@

```{r fig-widths}
fig_wid <- c(".3\\textwidth",
             ".6\\textwidth",
             ".8\\textwidth",
             "\\textwidth")
```

<!-- randomization -->
```{r sstat, fig.cap="Distribution of simulated and observed *S* statistic", out.width=fig_wid[1]}
knitr::include_graphics(here("plots/cooc_sstat.pdf"))
```

@@

```{r label-cooc-nets}
label_cooc_nets <- paste("Networks of variables frequently present together and excluding each other.",
                         "1 – bottle, 2 – bowl, 3 – globular pot, 4 – polished stone tool, 5 – local lithics,",
                         "6 – non-local lithics, 7 – pebble, 8 – grinding slab, 9 – bone tool,",
                         "10 – U-shaped pendant, 11 – spondylus beads, 12 – O-shaped buckle.",
                         "Images adapted from Podborský 2002")
```
```{r cooc-nets, fig.cap=label_cooc_nets, out.width=fig_wid[4]}
knitr::include_graphics(here::here("plots", "cooc_nets_fin.jpeg"))
```

<!-- plans -->
```{r plan-sex, fig.cap="Plan of the cemetery, kernel density estimation for differently sexed bodies", out.width=fig_wid[4]}
knitr::include_graphics(here("text/figs/plan_sex.pdf"))
```

```{r plan-age, fig.cap="Plan of the cemetery, kernel density estimation for bodies of different ages", out.width=fig_wid[4]}
knitr::include_graphics(here("text/figs/plan_age.pdf"))
```

```{r plan-orig, fig.cap="Plan of the cemetery, kernel density estimation for bodies of different origins", out.width=fig_wid[2]}
knitr::include_graphics(here("text/figs/plan_origin.pdf"))
```

@@

```{r label-pp-funs}
label_pp_funs <- paste(
  "Functions *G*, *F*, *K* and *T*.",
  "*Dashed line* -- theoretical curve under complete spatial randomness,
  *solid line* -- empirical curve, *gray area* -- envelope")
```
```{r pp-funs, fig.cap=label_pp_funs, out.width=fig_wid[4]}
knitr::include_graphics(here("text/figs/pp_funs.pdf"))
```

```{r label-pp-crossk}
label_crossk <- paste("Multitype K function for different categorical variables; for legend, see previous Figure")
```
```{r pp-crossk, fig.cap=label_crossk, out.width=fig_wid[3]}
knitr::include_graphics(here("text/figs/pp_crossk.pdf"))
```

<!-- neighborhood -->
```{r plan-graphs, fig.cap="Neighborhood graphs: Delaunay triangulation and Gabriel graph", out.width=fig_wid[4]}
knitr::include_graphics(here("text/figs/plan_graphs.pdf"))
```

@@

```{r label-perc-dist}
label_perc_dist <- "Normalized sizes of largest clusters at given distances, percolation thresholds labeled"
```
```{r perc-distance, fig.cap=label_perc_dist, out.width=fig_wid[2]}
knitr::include_graphics(here("text/figs/perc_distance.pdf"))
```

```{r perc-plan, fig.cap="Clusters formed at given percolation threshold distances", out.width=fig_wid[3]}
knitr::include_graphics(here("text/figs/perc_plan_ids.pdf"))
```

>>>>

The given statistics for the independent random process are estimated using Monte Carlo
simulation with 999 iterations.
Although a smaller number of permutations should be sufficient to achieve
credible results
[@baddeley2015, 384], we follow the approach taken by @sayer2013 [78].
The ten most extreme realizations of the simulation are discarded as outliers, so
the significance level is 0.01 [cf. @baddeley2015 410].
The remaining realizations of given point processes are used to construct envelopes
around function curves derived from observed data.

@@

<!-- software packages -->
Data is analyzed in an R environment for statistical computing and graphics
[@rcoreteam2021] using functions of several different packages; 
the most frequently used are packages from the *tidyverse* family [@wickham2019a].
Packages *sf* [@pebesma2018], *spatstat* [@baddeley2015], *spdep* [@bivand2018] and 
*concaveman* [@gombin2020] are used to perform various spatial analysis tasks; 
*igraph* [@csardi2006a] to construct and analyze networks;
*vegan* [@oksanen2020] and *randomizr* [@coppock2019] in the randomization experiments.

@@

```{r fig-widths}
fig_wid <- c(".3\\textwidth",
             ".6\\textwidth",
             ".8\\textwidth",
             ".98\\textwidth")
```

<!-- randomization -->
```{r sstat, fig.cap="Distribution of simulated and observed *S* statistic", out.width=fig_wid[1]}
knitr::include_graphics(here("plots", "cooc_sstat.pdf"))
```

@@

```{r label-cooc-nets}
label_cooc_nets <- paste("Networks of variables frequently present together and excluding each other.",
                         "1 – bottle, 2 – bowl, 3 – globular pot, 4 – polished stone tool, 5 – local lithics,",
                         "6 – non-local lithics, 7 – pebble, 8 – grinding slab, 9 – bone tool,",
                         "10 – U-shaped pendant, 11 – spondylus beads, 12 – O-shaped buckle.",
                         "Images adapted from Podborský 2002")
```
```{r cooc-nets, fig.cap=label_cooc_nets, out.width=fig_wid[4]}
knitr::include_graphics(here("plots", "cooc_nets_fin.jpeg"))
```

<!-- plans -->
```{r plan-sex, fig.cap="Plan of the cemetery, kernel density estimation for differently sexed bodies", out.width=fig_wid[4]}
knitr::include_graphics(here("plots", "plan_sex.pdf"))
```

```{r plan-age, fig.cap="Plan of the cemetery, kernel density estimation for bodies of different ages", out.width=fig_wid[4]}
knitr::include_graphics(here("plots", "plan_age.pdf"))
```

```{r plan-orig, fig.cap="Plan of the cemetery, kernel density estimation for bodies of different origins", out.width=fig_wid[2]}
knitr::include_graphics(here("plots", "plan_origin.pdf"))
```

@@

```{r label-pp-funs}
label_pp_funs <- paste(
  "Functions *G*, *F*, *K* and *T*.",
  "*Dashed line* - theoretical curve under complete spatial randomness;
  *solid line* - empirical curve; *gray area* - envelope")
```
```{r pp-funs, fig.cap=label_pp_funs, out.width=fig_wid[4]}
knitr::include_graphics(here("plots", "pp_funs.pdf"))
```

```{r label-pp-crossk}
label_crossk <- paste("Multitype K function for different categorical variables; for legend, see previous Figure")
```
```{r pp-crossk, fig.cap=label_crossk, out.width=fig_wid[3]}
knitr::include_graphics(here("plots", "pp_crossk.pdf"))
```

<!-- neighborhood -->
```{r plan-graphs, fig.cap="Neighborhood graphs: Delaunay triangulation and Gabriel graph", out.width=fig_wid[4]}
knitr::include_graphics(here("plots", "plan_graphs.pdf"))
```

@@

```{r label-perc-dist}
label_perc_dist <- "Normalized sizes of largest clusters at given distances, percolation thresholds labeled"
```
```{r perc-distance, fig.cap=label_perc_dist, out.width=fig_wid[2]}
knitr::include_graphics(here("plots", "perc_distance.pdf"))
```

```{r perc-plan, fig.cap="Clusters formed at given percolation threshold distances", out.width=fig_wid[3]}
knitr::include_graphics(here("plots", "perc_plan_ids.pdf"))
```
